---
title: 'Lab 2: Probability Theory'
output:
  html_document:
    df_print: paged
subtitle: 'R script for question 4e to 4i'
header-includes: \usepackage{amsmath} \newcommand{\var}{\mathrm{var}}
---


# 4. Circles, Random Samples, and the Central Limit Theorem

##e. Now let $n=100$.  Use R to simulate a draw for $X_1, X_2,...,X_n$ and $Y_1,Y_2,...,Y_n$.  Calculate the resulting values for $D_1, D_2,...D_n$.  Create a plot to visualize your draws, with $X$ on one axis and $Y$ on the other.  We suggest using a command like the following to assign a different color to each point, based on whether it falls inside the unit circle or outside it.  Note that we pass $d+1$ instead of $d$ into the color argument because 0 corresponds to the color white.

```{r}

# (Atit) Answer to question 4e:

# Set the number of simulation
n <- 100


# Generate random draws from a uniform distribution with range [-1,1]
# for both the r.v. X and the r.v. Y.
x <- runif(n, min=-1, max=1)
y <- runif(n, min=-1, max=1)


# Calculate the r.v. D.
d <- ifelse((x^2 + y^2) < 1, 1, 0)


# Plot X, Y, and D
plot(x,y, col=d+1, asp=1)
```



##f. What value do you get for the sample average, $\bar{D}$?  How does it compare to your answer for part a?

###(Atit) Answer to question  4f:

###The sample average, $\bar{D}$ is:

```{r}
# Calculate mean of D
mean(d)
```

###The expected value of D from 4a is pi/4:

```{r}
# Calculate mean of D
pi/4
```

###The value is a bit different.


##g. Now use R to replicate the previous experiment 10,000 times, generating a sample average of the $D_i$ each time.  Plot a histogram of the sample averages.

```{r}

#####################################################################
# Setup a function to generate a vector of D

simulate_d = function(n){
  
# Generate random draws from a uniform distribution with range [-1,1]
# for both the r.v. X and the r.v. Y.
x <- runif(n, min=-1, max=1)
y <- runif(n, min=-1, max=1)


# Calculate the r.v. D.
sim_d <- ifelse((x^2 + y^2) < 1, 1, 0)

return(sim_d)
}

rep_d <- replicate(100000,mean(simulate_d(n)))


# Plot the histogram of the simulated sample means
hist(rep_d, breaks = 50, main = "Simulated Sample Means from Repeated Sampling",
     xlab = "sample mean")
     
```


##h. Compute the standard deviation of your sample averages to see if it's close to the value you expect from part c.

###(Atit) Answer to question  4h:

```{r}
# Compute standard deviation of the sample averages:
sd(rep_d)
```

###This value is close the what is computed in 4c which was ~ 0.4105 / sqrt(100) = 0.04105.

##i. Compute the fraction of your sample averages that are larger that $3/4$ to see if it's close to the value you expect from part d.

###(Atit) Answer to question  4i:

```{r}
# Compute the fraction of thhe sample averages that are larger than 3/4:
sum(rep_d[rep_d > 0.75])/sum(rep_d)
```

###This value is close the what is computed in 4d which was 0.8057.
